Towards open collaboration in biomedical and health research: An open collaborative scoping review

### 2025-01-08 doc/paper.qmd
Update paper.qmd
Vomited some small sections of the results just so that we can discuss about the voice and structure.

>>>>

### Scoping review identifies only 11 papers discussing open collaboration

### Practices of open collaboration

Across the papers found in our literature search, X out of 11 discuss actual experiences on implementing open collaboration, albeit in different organizational levels. These range from lab level (Turoman et al) center level (Bush et al and Grange et al), multi-center level (ManyPrimates et al and Grange et al) and field level: psychology (Alessandorni et al), Spinal Cord injury (Torres-Espin et al) and neuroimaging (Niso et al). All of these share in common the need for an infrastructure that promotes open science and open collaboration via starting building a community of collaborators or making tools (usually online resources) that researchers can use to reach FAIR (Finndable, Accessible, Interoperable and Reusable) goals while doing research. Before delving in those, here we provide the specific needs for implementing open collaboration in each organizational level, as well as describing the differences in approaches among experiences between and across levels (Figure 2).

#### Implementation of open collaboration practices at laboratory level

#### Implementation of open collaboration practices at center level

For center-level only one paper was identified (Bush et al), providing the example of a center of neuroimaging describing first how they identified the needs for open collaboration, then the individual-level practices that researchers should adapt and then the center-level changes to promote reaching FAIR goals.

#### Implementation of open collaboration practices at multi-center level

Interestingly, the level of organization with most papers is the multi-center level with two papers from ManyPrimates et al (ManyPrimates 2019 and ManyPrimates 2021) and Grange et al. In this organization level the most urgent needs are: 1) identifying the needs for open collaboration and 2) creating an infrastructure that allows those needs to be satisfied. Indeed, Grange et al describes a systematic review as an exploration of the needs for 11 UKRN centers to implement open collaborative practives. In this paper they specially focus on tools and resources among the 11 institutions and make clear the need for starting to utilize open resources as a first step to adapt open collaborative practices. The first paper from ManyPrimates et al [REF] describes the start of another multi-center collaboration, albeit from different institutions. It is interesting to observe that in this situation, the complex issue is to start the collaboration itself, which was ignited thanks to a symposium and an email chain among future collaborators. In contrast, ManyPrimates could quickly develop an open collaborative structure with relative ease in comparison to changing an already solid infrastructure, like in the case of Grange et al.

#### Implementation of open collaboration practices at field level

### Useful tools for open collaboration
